---
title: "Advanced plotting"
output: ioslides_presentation
---
```{r}

library(ggplot2)
library(dplyr)

data("diamonds")

names(diamonds)
```
# Create a histogram of diamond prices.
# Facet the histogram by diamond color
# and use cut to color the histogram bars.
```{r}

ggplot(data= diamonds, aes(x= price))+
  geom_histogram(aes(color = cut, fill = cut))+ # could use position = "dodge" to set bars beside
  scale_x_log10()+
  scale_fill_brewer( type = 'qual' , palette = 'Set2')+
  facet_wrap(~ color)+
  theme_dark() # and a lot other theme with other option like font
```


# Create a scatterplot of diamond price vs.
# table and color the points by the cut of
# the diamond.
```{r}
ggplot(data = diamonds, aes(table, price))+
  geom_point(aes(color = cut), size = 3)+
  scale_fill_brewer( type = 'qual')+
  scale_x_continuous(breaks = seq(50,80,2))+
  coord_cartesian( xlim= c(50,80))
```



# Create a scatterplot of diamond price vs.
# volume (x * y * z) and color the points by
# the clarity of diamonds. Use scale on the y-axis
# to take the log10 of price. You should also
# omit the top 1% of diamond volumes from the plot.
```{r}

diamonds$volume <- diamonds$x * diamonds$y * diamonds$z

ggplot(data= diamonds, aes(volume, price))+
  geom_point(aes(color = clarity))+
  coord_cartesian(xlim= c(0, quantile(diamonds$volume, 0.99)) )+
  scale_color_brewer(type = 'div')+
  scale_y_log10()

```
# ____________________________________________________________________
```{r}

setwd('C:\\Users\\secretary-1\\Udacity\\Data Analysis\\P5')
pf <- read.delim('pseudo_facebook.tsv')

```

# Create a line graph of the median proportion of
# friendships initiated ('prop_initiated') vs.
# tenure and color the line segment by year_joined.bucket.
```{r}

pf$prop_initiated <- pf$friendships_initiated / pf$friend_count
pf$year_joined <- floor(2014 - pf$tenure / 365)
pf$year_joined.bucket <- cut(pf$year_joined, breaks = c(2004, 2009, 2011, 2012, 2014), labels = c("(2004,2009]", "(2009,2011]", "(2011,2012]", "(2012,2014]"))

ggplot(data = subset(pf, !is.na(tenure)), aes(30*(round(tenure/30)), prop_initiated))+
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)
```
```{r}
# another way
ggplot(data = subset(pf, !is.na(tenure)), aes(tenure, prop_initiated))+
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)+
  geom_smooth()

```


# mean of the 2012-2014 group after excluding na
```{r}
group_2012 <- subset(pf, year_joined.bucket == '(2012,2014]' & !is.na(prop_initiated))
mean(group_2012$prop_initiated)
```
```{r}
ggplot(data = subset(pf, !is.na(tenure)), aes(30*(round(tenure/30)), prop_initiated))+
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)+
  geom_hline(yintercept= mean(group_2012$prop_initiate))
```
     


# Create a scatter plot of the price/carat ratio
# of diamonds. The variable x should be
# assigned to cut. The points should be colored
# by diamond color, and the plot should be
# faceted by clarity.
```{r}
ggplot(data= diamonds, aes(cut, price/carat))+
  geom_jitter(aes(color = color))+
  facet_wrap(~clarity)+
  scale_color_brewer(type = 'div')
```


