Lesson 3
========================================================

***

### What to Do First?
Notes:
check the variable we have

***

### Pseudo-Facebook User Data
Notes:
read file and set the sep to be tab '\t'

```{r Pseudo-Facebook User Data}
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
library(ggplot2)

```

***

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
install.packages('ggplot2')
library(ggplot2)
qplot(x = dob_day, data = pf)
  
```

***

#### What are some things that you notice about this histogram?
Response:

***

### Moira's Investigation
Notes:

***

### Estimating Your Audience Size
Notes:

***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response:

#### How many of your friends do you think saw that post?
Response:

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response:

***

### Perceived Audience Size
Notes:

***
### Faceting
Notes:

```{r Faceting}
ggplot(data = pf , x = dob_day) +
  facet_wrap(~dob_day, ncol = 3)  # spilt data in several chart
  #facet_grid(vert col ~ horz col2)
```

#### Let’s take another look at our plot. What stands out to you here?
Response:

***

### Be Skeptical - Outliers and Anomalies
Notes:

***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response:

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
qplot(data = pf , x = friend_count, binwidth = 25, xlim = c(0, 1000))+
  facet_wrap(~gender)

```

#### How is this plot similar to Moira's first plot?
Response:

***

### Limiting the Axes
Notes:

```{r Limiting the Axes}

```

### Exploring with Bin Width
Notes:

***

### Adjusting the Bin Width
Notes:

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))
```

***

### Omitting NA Values
Notes:

```{r Omitting NA Values}

```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}
by(pf$friend_count, pf$gender, summary)
```

#### Who on average has more friends: men or women?
Response:

#### What's the difference between the median friend count for women and men?
Response:

#### Why would the median be a better measure than the mean?
Response:

***

### Tenure
Notes:

```{r Tenure}
qplot(data= pf, x= tenure/365, binwidth=.25, color=I('red'))+
  scale_x_continuous(breaks = seq(0, 7,0.2), limits = c(0,7))
```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}

```

***

### Labeling Plots
Notes:

```{r Labeling Plots}

```

***

### User Ages
Notes:

```{r User Ages}
qplot(data=pf, x= age, binwidth = 1)
```

#### What do you notice?
Response:

***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:

***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}
install.packages('gridExtra')
library(gridExtra)
```

***


### Frequency Polygons

```{r Frequency Polygons}
ggplot(aes(x = friend_count, y = ..count../sum(..count..)),
       data = subset(pf, !is.na(gender))) +
  geom_freqpoly(aes(color = gender), binwidth=10) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  xlab('Friend Count') +
  ylab('Proportion of users with that friend count')
```

***

### Likes on the Web
Notes:

```{r Likes on the Web}
qplot(data = subset(pf, !is.na(gender)), x = www_likes, geom = 'freqpoly', color = gender)+
  scale_x_continuous()+
  scale_x_log10()
by(pf$www_likes, pf$gender, sum)
```


***

### Box Plots
Notes:

```{r Box Plots}
qplot(data = subset(pf, !is.na(gender)), x = gender, y = friend_count, geom='boxplot')+
    scale_y_continuous(limits = c(0, 1000))

```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}

qplot(data = subset(pf, !is.na(gender)), x = gender, y = friend_count, geom='boxplot')+
    coord_cartesian(ylim = c(0, 500))
```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}
qplot(data = subset(pf, !is.na(gender)), x = gender, y = friendships_initiated, geom='boxplot')+
    coord_cartesian(ylim = c(0, 150))
```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}
summary(pf$mobile_likes >0)
pf$mobile_check_in <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes>0, 1, 0)
summary(pf$mobile_check_in)
pf$mobile_check_in <- factor(pf$mobile_check_in)
summary(pf$mobile_check_in)


```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}

```

Response:

***

### Analyzing One Variable
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!