Lesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
library(ggplot2)
library(tidyr)
library(reshape2)
library(dplyr)

pf <- read.delim('pseudo_facebook.tsv')

ggplot(aes(x = age), data = subset(pf, !is.na(gender)))+
  geom_histogram()+
  facet_wrap(~gender)
```

```{r Plotting Conditional Summaries}
ggplot(aes(x = gender, y= age), data = subset(pf, !is.na(gender)))+
  geom_boxplot()+
  stat_summary(fun.y = mean, geom = 'point', shape = 4)

```


```{r Plotting Conditional Summaries}
ggplot(aes(x = age, y= friend_count), data = subset(pf, !is.na(gender)))+
  geom_line(aes(color = gender), stat= 'summary', fun.y = median)

```

```{r Plotting Conditional Summaries}
pf.fc_by_age_gender <- pf %>%
  filter(!is.na(gender)) %>%
  group_by(age, gender) %>%
  summarise(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n = n())

head(pf.fc_by_age_gender,3)
```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
ggplot(aes(x = age, y= median_friend_count), data = pf.fc_by_age_gender)+
  geom_line(aes(color = gender))

```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
pf.fc_by_age_gender.wide <-
  subset(pf.fc_by_age_gender[c('age', 'gender', 'median_friend_count')],
         !is.na(gender)) %>%
  spread(gender, median_friend_count) %>%
  mutate(ratio = male / female)

head(pf.fc_by_age_gender.wide, 3)

```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
ggplot(data= pf.fc_by_age_gender.wide, aes(age, male))+
  geom_line(color = 'green', linetype = 5)+
  geom_line(aes(age, female), color = 'red', linetype = 4)+
  geom_hline(yintercept= mean(pf.fc_by_age_gender.wide$male), color='red', linetype = 3)+
  geom_hline(yintercept= mean(pf.fc_by_age_gender.wide$female), color='green', linetype = 5)

```

```{r}
ggplot(data=pf.fc_by_age_gender.wide, aes(age, female/male))+ # checking the ration btw f & m
  geom_line()+
  geom_hline(yintercept = 1, linetype=2 , alpha=0.5) # intercept at 1 as its ration when f=m

```
***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
pf$year_joined <- floor(2014 - pf$tenure / 365)

```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
pf$year_joined.bucket <- cut(pf$year_joined, breaks = c(2004, 2009, 2011, 2012, 2014), labels = c("(2004,2009]", "(2009,2011]", "(2011,2012]", "(2012,2014]"))

table(pf$year_joined.bucket)
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
ggplot(data= subset(pf, !is.na(year_joined.bucket)), aes(age, friend_count, color= year_joined.bucket))+
  geom_line(stat = 'summary', fun.y = median)+
  coord_cartesian(xlim= c(min(pf$age),90))
```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
mean_data <- subset(pf, !is.na(age))

ggplot(data= subset(pf, !is.na(year_joined.bucket)), aes(age, friend_count))+
  geom_line(aes(color = year_joined.bucket),stat = 'summary', fun.y = mean)+
  coord_cartesian(xlim= c(min(pf$age),90))+
  geom_line(stat = 'summary', fun.y = mean, linetype = 2)
```

***

### Friending Rate
Notes:

```{r Friending Rate}
friend_tenure <- subset(pf, tenure>1)
friend_tenure$rate <- friend_tenure$friend_count / friend_tenure$tenure

summary(friend_tenure$rate)

# alternative
with(subset(pf, tenure>1), summary(friend_count/tenure))
```

***

### Friendships Initiated
Notes:

What is the median friend rate?

What is the maximum friend rate?

```{r Friendships Initiated}
ggplot(data= subset(pf, tenure>0), aes(tenure, friendships_initiated/tenure))+
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean)+
  scale_x_continuous(breaks = seq(0,3000,500))

```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure), # every day
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure), # count only week or more
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure), # count month or more
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x =tenure, y = friendships_initiated / tenure), # count 3 month or more
       data = subset(pf, tenure > 0)) +
  geom_smooth(aes(color = year_joined.bucket))

```
```{r}
pf$tenure7 <- 7 * round(pf$tenure/7)
pf$tenure30 <- 30 * round(pf$tenure/30)
pf$tenure90 <- 90 * round(pf$tenure/90)

View(pf,)
```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}

```

***

### Number of Purchases
Notes:

```{r Number of Purchases}

```

***

### Prices over Time
Notes:

```{r Prices over Time}

```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}

```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

