Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)

pf <- read.delim('pseudo_facebook.tsv')

qplot(data= pf, x= age, y= friend_count)

```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
ggplot(aes(x = age, y = friend_count), data = pf, xlim= c(13,90)) +
  geom_point()
```

***

### Overplotting
Notes:

```{r Overplotting}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_jitter(alpha=1/20)+
  xlim(13,90)
```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_jitter(alpha=1/20)+
  xlim(13,90)
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point(alpha=1/20)+
  xlim(13,90)+
  coord_trans(y='sqrt')
```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
ggplot(aes(x = age, y = friendships_initiated), data = pf) +
  geom_jitter(alpha = 0.05, position = position_jitter(h=0))+
  xlim(13,90)+
  coord_trans(y='sqrt')

```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
library(dplyr)

age_groups <- group_by(pf, age)
pf.fc_by_age <- summarise(age_groups, 
          friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n = n())
arrange(pf.fc_by_age, age) # sort by age

```

Create your plot!

```{r Conditional Means Plot}
pf_fc_by_age <- pf %>%
  group_by(age) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %>%
  arrange(age)

head(pf_fc_by_age, 20)
```

```{r}
ggplot(data= pf_fc_by_age, aes(age, friend_count_mean, color = friend_count_mean))+
  geom_line(size = 1)
```


***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point(alpha = 0.05, position = position_jitter(h=0), color = 'orange')+
  coord_cartesian(xlim = c(13, 90))+
  coord_trans(y='sqrt')+
  geom_line(stat = 'summary', fun.y = mean)+
  geom_line(stat = 'summary', fun.y = median, color= 'red')+
  geom_line(stat= 'summary', fun.y = quantile, fun.args = list(probs = .1), linetype= 2, color= 'blue')
ggsave('save.jpeg')
```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
cor.test(pf$age, pf$friend_count)
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(pf, age <= 70), cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}
ggplot(aes(www_likes_received, likes_received), data= pf)+
  coord_cartesian(xlim=c(0, 10000), ylim = c(0,10000))+
  geom_point(alpha = 0.05, position = position_jitter(h=0))
  
  
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
ggplot(aes(www_likes_received, likes_received), data= pf)+
  coord_cartesian(xlim=c(0, quantile(pf$www_likes_received, 0.95)),
                  ylim = c(0, quantile(pf$likes_received, 0.95)))+
  geom_point(alpha = 0.05, position = position_jitter(h=0))+
  geom_smooth(method = 'lm', color = 'red')
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
with(pf, cor.test(www_likes_received, likes_received))
```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)

data("Mitchell")
```

Create your plot!

```{r Temp vs Month}
ggplot(data= Mitchell, aes(x= Month, y = Temp))+
  geom_point()
  
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
cor.test(Mitchell$Month, Mitchell$Temp)
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
ggplot(data= Mitchell, aes(x= Month%%12, y = Temp))+
  geom_point()+
  scale_x_continuous(breaks = seq(0,204, 12))
```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
pf$age_with_months <- pf$age+((12-pf$dob_month)/12)

```

***

### Age with Months Means

```{r Age with Months Means}
pf.fc_by_age_months <- pf %>%
  group_by(age_with_months) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %>%
  arrange(age_with_months)

head(pf.fc_by_age_months)
```

Programming Assignment
```{r Programming Assignment}
ggplot(data= subset(pf.fc_by_age_months,age_with_months<71) , aes(age_with_months, friend_count_mean))+
  geom_line()
```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}

```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}
p1 <- ggplot(data= subset(pf.fc_by_age_months,age_with_months<71) , aes(age_with_months, friend_count_mean))+
  geom_line()+
  geom_smooth()

p2 <- ggplot(data= subset(pf_fc_by_age, age<71), aes(age, friend_count_mean))+
  geom_line()+
    geom_smooth()

p3 <- ggplot(data= subset(pf, age<71), aes(round(age/5)*5, friend_count))+
  geom_line(stat = 'summary', fun.y = mean)

library(gridExtra)
grid.arrange(p1, p2, p3, ncol = 1)
```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

