```{r, echo=FALSE}
knitr::opts_chunk$set(error = TRUE)
```
Red Wine dataset analysis by Ahmed Shaaban
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.
library(ggplot2)
library(gridExtra)
```
> **Introduction to this dataset:** 

- This dataset is public available for research. The details are described in [Cortez et al., 2009]
- Modeling wine preferences by data mining from physicochemical properties.
- The inputs include objective tests and the output is based on sensory data (median of at least 3 evaluations made by wine experts). Each expert graded the wine quality between 0 (very bad) and 10 (very excellent).
- Number of Instances 1599, Number of Attributes 11 + the quality graded.

** Our data set contains:**
```{r echo=FALSE, Load_the_Data}
# Load the Data
pf <-  read.csv('wineQualityReds.csv', row.names = 1)
str(pf)
```

# Univariate Plots Section


**Display data summary**
```{r echo=FALSE, Summary}
summary(pf)
```

**Notice:**
- I can see large differences between 3rd Qu. and max. in residual.sugar, free.sulfur.dioxide, total.sulfur.dioxide and sulphates which may contains outliers and will produce long tailled histogram.
- Quality grades in between 3 and 8.


**Display all the attributes distribution **
```{r echo=FALSE, Univariate_Plots}
p1 <- ggplot(pf, aes(fixed.acidity))+
  geom_histogram(color = 'black', fill = '#ff4d4d', binwidth = 0.5)

p2 <- ggplot(pf, aes(volatile.acidity))+
  geom_histogram(color = 'black', fill = '#ff4d4d', binwidth = 0.05)

p3 <- ggplot(pf, aes(citric.acid))+
  geom_histogram(color = 'black', fill = '#ff4d4d', binwidth = 0.03)

p4 <- ggplot(pf, aes(residual.sugar))+
  geom_histogram(color = 'black', fill = '#ff4d4d', binwidth = 0.3)

p5 <- ggplot(pf, aes(chlorides))+
  geom_histogram(color = 'black', fill = '#ff4d4d', binwidth = 0.02)

p6 <- ggplot(pf, aes(free.sulfur.dioxide))+
  geom_histogram(color = 'black', fill = '#ff4d4d', binwidth = 1.5)

p7 <- ggplot(pf, aes(total.sulfur.dioxide))+
  geom_histogram(color = 'black', fill = '#ff4d4d', binwidth = 7)

p8 <- ggplot(pf, aes(density))+
  geom_histogram(color = 'black', fill = '#ff4d4d', binwidth = 0.001)

p9 <- ggplot(pf, aes(pH))+
  geom_histogram(color = 'black', fill = '#ff4d4d', binwidth = 0.04)

p10 <- ggplot(pf, aes(sulphates))+
  geom_histogram(color = 'black', fill = '#ff4d4d', binwidth = 0.05)

p11 <- ggplot(pf, aes(alcohol))+
  geom_histogram(color = 'black', fill = '#ff4d4d', binwidth = 0.2)

grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11)
```

**According to the attributes distribution above, all attributes seems right skewed with long right tail especially residual sugar, free sulfur dioxide and total sulfur dioxide, except pH and density are asymmetrically distributed **


**Display the quailty grades**
```{r echo=FALSE, Univariate_Plots_1}
ggplot(pf, aes(quality))+
  geom_bar(color = 'black', fill = '#ff4d4d')+
  coord_polar()
```

**We can see that most of the quality rating given were 5 and 6**

# Univariate Analysis

### The structure of this dataset?

1. Data with 1599 tests (row) and 11 attributes (columns) + 1 extra column for the quailty grades between 0 & 10.

2. Attribute informations: 
- fixed acidity: most acids involved with wine or fixed or nonvolatile (do not evaporate readily)
- volatile acidity: the amount of acetic acid in wine (too high can lead bad vinegar taste)
- citric acid: found in small quantities, citric acid can add 'freshness' and flavor to wines
- residual sugar: Sweetness in wine, in our dataset between 0.90 - 15.9 
- chlorides: the amount of salt in the wine
- free sulfur dioxide: the free form of SO2 (it prevents microbial growth & the oxidation of wine)
- total sulfur dioxide: amount of free and bound forms of S02.
- density: the density of water is close to that of water depending on the percent alcohol and sugar content.
- pH: describes how acidic or basic a wine our data are between 2.74 - 4.014 on the pH scale.
- sulphates: a wine additive which can contribute to sulfur dioxide gas (S02) levels.
- alcohol: the percent alcohol content of the wine, in our dataset between 8.40 - 14.9

### What are the main features of interest in our dataset?

**First: Analysing quality is our aim**

**Second: Display correlation table to decide the interested features**
```{r echo=FALSE, Univariate_Plots_2}
round(cor(pf), 3)
```

1. Alcohol level has good impact on quality.
2. density aslo could affect the alcohol.
3. fixed acidity has strong positive relation with density.
4. pH has strong negitive relation with fixed acidity & citric.acid.


### Other feature in the dataset that will help support our investigation:
- As per the attribution description of density: "density of water is close to that of water depending on the percent alcohol and sugar content" so, I will consider residual sugar in the investigation.
- Strong relationship has been found between free.sulfur.dioxide and total.sulfur.dioxide which is logical.

### Creating new variables from existing variables in the dataset:
- According to our interests in the above features, new variable will be created to support our investigation:

1. since most of gardes given in 5 & 6 and maximum rate is 8, i will set defferent gardes for each one as follow: poor<=4, med_quality=5, good=6, v.good>=7. 
```{r echo=FALSE, Univariate_Plots_3}
pf$grades <- cut(pf$quality, breaks = c(2, 4, 5, 6, 8), labels = c("Poor", "Med_quality", "Good", "V.Good"))

table(pf$grades)
```

2. alcohol levels will be set as per summary as follow: low<= 9.5, low<Medium<11.5, High>11.5
```{r echo=FALSE, Univariate_Plots_4}
pf$Alcohol_precent <- cut(pf$alcohol, breaks = c(8, 9.5, 11.5,15), labels = c("Low", "Medium", "High"))

table(pf$Alcohol_precent)
```

3. The Sweetness is very important measure of wine taste and it will be set as follow: low_sweet<= 9.5, low<Medium_sweet<11.5, High_sweet>11.5

```{r echo=FALSE, Univariate_Plots_5}
pf$Sweetness <- cut(pf$residual.sugar, breaks = c(0.8, 2, 4,15.5), labels = c("Low_sweet", "Medium_sweet", "High_sweet"))

table(pf$Sweetness)
```

### Other investigations : 
- Density were presented with cm and should be converted into dm to match the other attributes.


# Bivariate Plots Section

```{r echo=FALSE, Bivariate_Plots_1}
ggplot(pf, aes(density, fixed.acidity, color = '#ff4d4d'))+
  geom_point(position = 'Jitter')+
  ggtitle('Relationship between Density and fixed acidity')

```

```{r echo=FALSE, Bivariate_Plots_2}
ggplot(pf, aes(density, color = grades))+
  geom_density()+
  ggtitle('Density shapes as per Quality grades')

```

```{r echo=FALSE, Bivariate_Plots_3}
ggplot(pf, aes(grades, fill = Alcohol_precent))+
  geom_bar()+
  ggtitle('Relationship between Quality and Alcohol level')

```


```{r echo=FALSE, Bivariate_Plots_4}
ggplot(pf, aes(citric.acid, pH, color = '#ff4d4d' ))+
  geom_point(position = 'jitter')+
  ggtitle('Relatioship between pH and citric acid')

```


```{r echo=FALSE, Bivariate_Plots_5}
ggplot(pf, aes(free.sulfur.dioxide, total.sulfur.dioxide))+
  geom_line()+
  facet_wrap(~ grades, ncol = 1)+
  ggtitle('Distribution of Total & Free sulfur.dioxide according to quality')

```


```{r echo=FALSE, Bivariate_Plots_6}
ggplot(pf, aes(alcohol, residual.sugar))+
  geom_line()+
  ggtitle('Relationship Alcohol and Residual sugar')

```


```{r echo=FALSE, Bivariate_Plots_7}
ggplot(pf, aes(chlorides, sulphates))+
  geom_point(alpha = 0.5, size = 2, position = 'Jitter')+
  ggtitle('Relationship between Chlorides and Sulphates')

```

```{r echo=FALSE, Bivariate_Plots_8}
ggplot(pf, aes(volatile.acidity, fixed.acidity))+
  geom_point(alpha = 0.5, size = 2, position = 'Jitter')+
  ggtitle('Relationship between volatile acidity and fixed acidity')

```

# Bivariate Analysis


### Relationships I observed in the above plots:

1. Strong positive relationship between density and fixed acidity but we need to convert density into dm^3.
2. Comparing density shapes according to quality grades does not give much information, but I found that at low density (0.990 to 0.991) we received only good and very good quality rating.
3. Relationship between alcohol level and quality grades shows that almost all the very good rates have been given to high & medium alcohol level, and the same to good rates.
4. Relatioship between pH and citric acid shows weak negtive relationship.
5. Logical relationship between free & total sulfur dioxide, but when we facet it to quality grades we found strange behavior, and we need to investigate more.
6. Relationship Alcohol and Residual sugar does not show much information but I can see that Residual sugar getting a little bit lower as alcohol increase.
7. Relationship between Chlorides and Sulphates shows non linear relationship, and indicate positive relationship. 
8. Relationship between volatile.acidity and fixed.acidity shows also non linear relationship does not give information. 

# Multivariate Plots Section

## Plot one:
```{r echo=FALSE, Multivariate_Plots_1}
ggplot(subset(pf, Alcohol_precent == 'High'), 
       aes(residual.sugar, color = Sweetness, fill = Sweetness))+
  geom_bar()+
  scale_y_log10()+
  ylab('Transforming log(frequency)')+
  facet_wrap(~grades, ncol = 4)+
  ggtitle('Relationship between Quality and Sweetness at high alcohol level')
```

## Plot Two:
```{r echo=FALSE, Multivariate_Plots_2}
 ggplot(pf, aes(chlorides, sulphates,color = grades))+
  geom_point(alpha = 0.5, size = 2, position = 'Jitter')+
  scale_x_continuous(breaks = seq(0,11,0.02))+
  scale_y_continuous(breaks = seq(0, 1.25,0.2))+
  coord_cartesian(xlim = c(min(pf$chlorides), quantile(pf$chlorides, 0.98)), 
                  ylim = c(min(pf$sulphates), quantile(pf$sulphates, 0.99)))+
  scale_color_brewer(type = 'div', 
                     guide = guide_legend(title = 'Quality Grades', reverse = T, 
                                          override.aes = list(alpha = 1, size = 3)))+
  geom_hline(yintercept= c(0.65, 1.05), color='black', linetype = 2)+
  geom_vline(xintercept= c(0.03, 0.1), color='black', linetype = 2)+
  xlab('sodium chloride - g / dm^3')+
  ylab('potassium sulphate - g / dm3')+
  ggtitle('Relationship between Chlorides and Sulphates according to quality')

```

**Omitting 1% of top sulphates values and 2% of top chlorides values**


# Multivariate Analysis

### Very interesting relationship relationships I observed:

1. Since we observed some relation between alcohol and residual sugar in the bivariate analysis section as the residual.sugar getting a little bit lower at high alcohol, I intended to look more closely by subsetting our data by high alcohol level and get the effect of quality grades.

2. Relationship between chlorides and sulphates is positive but we need to look closely at the quality grades applied to it.

### Were there any interesting or surprising interactions between features?

1. Relation between alcohol and residual sugar: I found that medium sweet and low sweet wine has much better rating than high sweet wine at the high level alcohol(high alcohol level receive better rating as we mentioned before).

2. Relation between chlorides and sulphates: We could find the best wine rating when chlorides is between 0.03-0.1 and sulphates between 0.65-1.05, as we found most of Very good and good rating in this range along with some medium quality rates.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}
ggplot(pf, aes(free.sulfur.dioxide, total.sulfur.dioxide, color = grades))+
  geom_line(stat = 'summary', fun.y = mean)+
  coord_cartesian(xlim = c(min(pf$total.sulfur.dioxide), quantile(pf$free.sulfur.dioxide, 0.99)),
                  ylim = c(min(pf$total.sulfur.dioxide), quantile(pf$total.sulfur.dioxide, 0.99)))+
  ylab('total sulfur dioxide (mg/dm^3)')+
  xlab('free sulfur dioxide (mg/dm^3)')+
  ggtitle('Relationship between Average Total & Free sulfur dioxide coloured by Quality grades')
```

**Omitting 1% of top total & free sulfur.dioxide**

### Description One
- Relationship between Total & Free sulfur dioxide has been presented by average mean, I choose this plot because all the poor rating wine with free sulfur.dioxide of 40 and after 40 rating is meduim quality, good and very good.
- After (free sulfur.dioxide of 40) less total sulfur dioxide means better quality rating.

### Plot Two
```{r echo=FALSE, Plot_Two}
ggplot(pf, aes(density/10, fixed.acidity, color = Alcohol_precent))+
  geom_point(alpha = 0.4, position = 'jitter')+
  scale_color_brewer(type = 'qual', palette = 'Set1',
                     guide = guide_legend(title = 'Alcohol_precent', reverse = T,
                                          override.aes = list(alpha = 1, size = 3)))+
  coord_cartesian(ylim = c(min(pf$fixed.acidity), quantile(pf$fixed.acidity, 0.99)))+
  facet_wrap(~grades, ncol = 4)+
  ylab('Fixed acidity - g / dm^3')+
  xlab('Density by dm (g / dm^3)')+
  ggtitle('Grades scatter plot by Density and fixed acidity colored by Alcohol precentage')
```

### Description Two
- Very interesting relationship between density and fixed.acidity as it is strong positive relationship, after converting density to dm and facet it to quality rating and coloured by alcohol levels I could not found evdince how fixed acidity and  density affect the quality but i found that best rates were given to high and medium alcohol llevels.

### Plot Three
```{r echo=FALSE, Plot_Three}
ggplot(pf, aes(citric.acid, pH, color = grades))+
  geom_point(alpha = 0.5, size = 2, position = 'jitter')+
  scale_color_brewer(type = 'div', 
                     guide = guide_legend(title = 'Quality Grades', reverse = T,
                                          override.aes = list(alpha = 1, size = 3)))+
  ylab('pH')+
  xlab('Citric acid (g / dm^3)')+
  ggtitle('Relatioship between pH and citric acid coloured by Quality')

```

### Description Three
- Interesting plot in our search of quality, relationship between pH and citric is negitive and weak but I can see that with low pH and high citric acid we can find good and very good quality.
------

# Reflection

> Red wine dataset needed a lot exploration to determin which attribute will affect the quality.
> Very interesting analysis that could help a lot to put the best standards for good wine.
> Many plots does not help a lot to determin quality but others surprised me such as the sweetness should be low or medium for best quality.