Red Wine dataset analysis by Ahmed Shaaban
========================================================

> **Tip**: You will see quoted sections like this throughout the template to
help you construct your report. Make sure that you remove these notes before
you finish and submit your project!

> **Tip**: One of the requirements of this project is that your code follows
good formatting techniques, including limiting your lines to 80 characters or
less. If you're using RStudio, go into Preferences \> Code \> Display to set up
a margin line to help you keep track of this guideline!

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(dplyr)
library(gridExtra)
library(reshape2)
library(gridExtra)
```
> **Introduction to this dataset:** 
- This dataset is public available for research. The details are described in [Cortez et al., 2009]
- Modeling wine preferences by data mining from physicochemical properties.
- The inputs include objective tests and the output is based on sensory data (median of at least 3 evaluations made by wine experts). Each expert graded the wine quality between 0 (very bad) and 10 (very excellent).
- Number of Instances 1599, Number of Attributes 11 + the quality graded.

```{r echo=FALSE, Load_the_Data}
# Load the Data
pf <-  read.csv('wineQualityReds.csv', row.names = 1)
```

> **Tip**: Before you create any plots, it is a good idea to provide a short
introduction into the dataset that you are planning to explore. Replace this
quoted text with that general information!


# Univariate Plots Section

> **Tip**: In this section, you should perform some preliminary exploration of
your dataset. Run some summaries of the data and create univariate plots to
understand the structure of the individual variables in your dataset. Don't
forget to add a comment after each plot or closely-related group of plots!
There should be multiple code chunks and text sections; the first one below is
just to help you get started.


**Display data summary**
```{r echo=FALSE, Load_the_Data}
summary(pf)
```
**Notice:**
- I can see large differences between 3rd Qu. and max. in residual.sugar, free.sulfur.dioxide, total.sulfur.dioxide and sulphates which may contains outliers and will produce long tailled histogram.
- Quality grades in between 3 and 8.


**Display all the attributes distribution **
```{r echo=FALSE, Univariate_Plots}
p1 <- ggplot(pf, aes(fixed.acidity))+
  geom_histogram(color = 'black', fill = '#ff4d4d', binwidth = 0.5)

p2 <- ggplot(pf, aes(volatile.acidity))+
  geom_histogram(color = 'black', fill = '#ff4d4d', binwidth = 0.05)

p3 <- ggplot(pf, aes(citric.acid))+
  geom_histogram(color = 'black', fill = '#ff4d4d', binwidth = 0.03)

p4 <- ggplot(pf, aes(residual.sugar))+
  geom_histogram(color = 'black', fill = '#ff4d4d', binwidth = 0.3)

p5 <- ggplot(pf, aes(chlorides))+
  geom_histogram(color = 'black', fill = '#ff4d4d', binwidth = 0.02)

p6 <- ggplot(pf, aes(free.sulfur.dioxide))+
  geom_histogram(color = 'black', fill = '#ff4d4d', binwidth = 1.5)

p7 <- ggplot(pf, aes(total.sulfur.dioxide))+
  geom_histogram(color = 'black', fill = '#ff4d4d', binwidth = 7)

p8 <- ggplot(pf, aes(density))+
  geom_histogram(color = 'black', fill = '#ff4d4d', binwidth = 0.001)

p9 <- ggplot(pf, aes(pH))+
  geom_histogram(color = 'black', fill = '#ff4d4d', binwidth = 0.04)

p10 <- ggplot(pf, aes(sulphates))+
  geom_histogram(color = 'black', fill = '#ff4d4d', binwidth = 0.05)

p11 <- ggplot(pf, aes(alcohol))+
  geom_histogram(color = 'black', fill = '#ff4d4d', binwidth = 0.2)

grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11)
```

**According to the attributes distribution above, all attributes seems right skewed with long right tail especially residual sugar, free sulfur dioxide and total sulfur dioxide, except pH and density are asymmetrically distributed **


**Display the quailty grades**
```{r echo=FALSE, Univariate_Plots}
ggplot(pf, aes(quality))+
  geom_bar(color = 'black', fill = '#ff4d4d')+
  coord_polar()
```

**We can see that most of the quality rating given were 5 and 6**

> **Tip**: Make sure that you leave a blank line between the start / end of
each code block and the end / start of your Markdown text so that it is
formatted nicely in the knitted text. Note as well that text on consecutive
lines is treated as a single space. Make sure you have a blank line between
your paragraphs so that they too are formatted for easy readability.

# Univariate Analysis

> **Tip**: Now that you've completed your univariate explorations, it's time to
reflect on and summarize what you've found. Use the questions below to help you
gather your observations and add your own if you have other thoughts!

### The structure of this dataset?
1. Data with 1599 tests (row) and 11 attributes (columns) + 1 extra column for the quailty grades between 0 & 10.
2. Attribute informations: 
- fixed acidity: most acids involved with wine or fixed or nonvolatile (do not evaporate readily)
- volatile acidity: the amount of acetic acid in wine (too high can lead bad vinegar taste)
- citric acid: found in small quantities, citric acid can add 'freshness' and flavor to wines
- residual sugar: Sweetness in wine, in our dataset between 0.90 - 15.9 
- chlorides: the amount of salt in the wine
- free sulfur dioxide: the free form of SO2 (it prevents microbial growth & the oxidation of wine)
- total sulfur dioxide: amount of free and bound forms of S02.
- density: the density of water is close to that of water depending on the percent alcohol and sugar content.
- pH: describes how acidic or basic a wine our data are between 2.74 - 4.014 on the pH scale.
- sulphates: a wine additive which can contribute to sulfur dioxide gas (S02) levels.
- alcohol: the percent alcohol content of the wine, in our dataset between 8.40 - 14.9

### What are the main features of interest in our dataset?

**First: Analysing quality is our aim**
**Second: Display correlation table to decide the interested features**
```{r echo=FALSE, Univariate_Plots}
round(cor(pf), 3)
```

1. Alcohol level has good impact on quality.
2. density aslo could affect the alcohol.
3. fixed acidity has strong positive relation with density.
4. pH has strong negitive relation with fixed acidity & citric.acid.


### Other feature in the dataset that will help support our investigation:
- As per the attribution description of density: "density of water is close to that of water depending on the percent alcohol and sugar content" so, I will consider residual sugar in the investigation.
- Strong relationship has been found between free.sulfur.dioxide and total.sulfur.dioxide which is logical.

### Creating new variables from existing variables in the dataset:
- According to our interests in the above features, new variable will be created to support our investigation:
1. since most of gardes given in 5 & 6 and maximum rate is 8, i will set defferent gardes for each one as follow: poor<=4, med_quality=5, good=6, v.good>=7. 
```{r echo=FALSE, Univariate_Plots}

pf$grades <- cut(pf$quality, breaks = c(2, 4, 5, 6, 8), labels = c("Poor", "Med_quality", "Good", "V.Good"))

table(pf$grades)
```

2. alcohol levels will be set as per summary as follow: low<= 9.5, low<Medium<11.5, High>11.5
```{r echo=FALSE, Univariate_Plots}

pf$Alcohol_precent <- cut(pf$alcohol, breaks = c(8, 9.5, 11.5,15), labels = c("Low", "Medium", "High"))

table(pf$Alcohol_precent)
```

3. The Sweetness is very important measure of wine taste and it will be set as follow: low_sweet<= 9.5, low<Medium_sweet<11.5, High_sweet>11.5

```{r echo=FALSE, Univariate_Plots}

pf$Sweetness <- cut(pf$residual.sugar, breaks = c(0.8, 2, 4,15.5), labels = c("Low_sweet", "Medium_sweet", "High_sweet"))

table(pf$Sweetness)
```

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?


# Bivariate Plots Section

> **Tip**: Based on what you saw in the univariate plots, what relationships
between variables might be interesting to look at in this section? Don't limit
yourself to relationships between a main output feature and one of the
supporting variables. Try to look at relationships between supporting variables
as well.

```{r echo=FALSE, Bivariate_Plots}
ggplot(pf, aes(density, fixed.acidity, color = '#ff4d4d'))+
  geom_point(position = 'Jitter')+
  ggtitle('Relationship between Density and fixed acidity')

```

```{r echo=FALSE, Bivariate_Plots}
ggplot(pf, aes(density, color = grades))+
  geom_density()+
  ggtitle('Density shapes as per Quality grades')

```
```{r echo=FALSE, Bivariate_Plots}
ggplot(pf, aes(grades, fill = Alcohol_precent))+
  geom_bar()+
  ggtitle('Relationship between Quality and Alcohol level')

```


```{r echo=FALSE, Bivariate_Plots}
ggplot(pf, aes(citric.acid, pH, color = '#ff4d4d' ))+
  geom_point(position = 'jitter')+
  ggtitle('Relatioship between pH and citric acid')

```


```{r echo=FALSE, Bivariate_Plots}
ggplot(pf, aes(free.sulfur.dioxide, total.sulfur.dioxide))+
  geom_line()+
  facet_wrap(~ grades, ncol = 1)
  ggtitle('Distribution of Total & Free sulfur.dioxide')

```

```{r echo=FALSE, Bivariate_Plots}
ggplot(pf, aes(alcohol, residual.sugar))+
  geom_line()+
  ggtitle('Relationship Alcohol and Residual sugar')

```


```{r echo=FALSE, Bivariate_Plots}
ggplot(pf, aes(chlorides, sulphates))+
  geom_point(alpha = 0.5, size = 2, position = 'Jitter')+
  ggtitle('Relationship between Chlorides and Sulphates')

```


# Bivariate Analysis

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.

### Relationships I observed in the above plots:

1. Strong positive relationship between density and fixed acidity.
2. Comparing density shapes according to quality grades does not give much information, but I found that at low density (0.990 to 0.991) we received only good and very good quality rating.
3. Relationship between alcohol level and quality grades shows that almost all the very good rates have been given to high & medium alcohol level, and the same to good rates.
4. Relatioship between pH and citric acid shows weak negtive relationship.
5. Logical relationship between free & total sulfur dioxide, but when we facet it to quality grades we found strange behavior, and we need to investigate more.
6. Relationship Alcohol and Residual sugar does not show much information but I can see that Residual sugar getting a little bit lower as alcohol increase.
7. Relationship between Chlorides and Sulphates shows non linear relationship, and indicate positive relationship. 
8. 
# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}
ggplot(subset(pf, Alcohol_precent == 'High'), 
       aes(residual.sugar, color = Sweetness, fill = Sweetness))+
  geom_bar()+
  scale_y_log10()+
  ylab('Transforming log(frequency)')+
  facet_wrap(~grades, ncol = 4)+
  ggtitle('Relationship between Quality and Sweetness at high alcohol level')
```

```{r echo=FALSE, Multivariate_Plots}
 ggplot(pf, aes(chlorides, sulphates,color = grades))+
  geom_point(alpha = 0.5, size = 2, position = 'Jitter')+
  scale_x_continuous(breaks = seq(0,11,0.02))+
  scale_y_continuous(breaks = seq(0, 1.25,0.2))+
  coord_cartesian(xlim = c(min(pf$chlorides), quantile(pf$chlorides, 0.98)), 
                  ylim = c(min(pf$sulphates), quantile(pf$sulphates, 0.99)))+
  scale_color_brewer(type = 'div', 
                     guide = guide_legend(title = 'Quality Grades', reverse = T, 
                                          override.aes = list(alpha = 1, size = 3)))+
  geom_hline(yintercept= c(0.65, 1.05), color='black', linetype = 2)+
  geom_vline(xintercept= c(0.03, 0.1), color='black', linetype = 2)+
  ggtitle('Relationship between Chlorides and Sulphates according to quality')

```
**Omitting 1% of top sulphates values and 2% of top chlorides values**

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

1. Since we observed some relation between alcohol and residual sugar in the bivariate analysis section as the residual.sugar getting a little bit lower at high alcohol, I intended to look more closely by subsetting our data by high alcohol level and get the effect of quality grades.

2. Relationship between chlorides and sulphates is positive but we need to look closely at the quality grades applied to it.

### Were there any interesting or surprising interactions between features?

1. Relation between alcohol and residual sugar: I found that medium sweet and low sweet wine has much better rating than high sweet wine at the high level alcohol(high alcohol level receive better rating as we mentioned before).

2. Relation between chlorides and sulphates: We could find the best wine rating when chlorides is between 0.03-0.1 and sulphates between 0.65-1.05, as we found most of Very good and good rating in this range along with some medium quality rates.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}
ggplot(pf, aes(free.sulfur.dioxide, total.sulfur.dioxide, color = grades))+
  geom_line(stat = 'summary', fun.y = mean)+
  coord_cartesian(xlim = c(min(pf$total.sulfur.dioxide), quantile(pf$free.sulfur.dioxide, 0.99)),
                  ylim = c(min(pf$total.sulfur.dioxide), quantile(pf$total.sulfur.dioxide, 0.99)))+
  ylab('Transforming log(frequency)')+
  ggtitle('Relationship between Total & Free sulfur.dioxide coloured by Quality grades')
```
**Omitting 1% of top total & free sulfur.dioxide**

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}
ggplot(pf, aes(density, fixed.acidity, color = Alcohol_precent))+
  geom_point(alpha = 0.3, position = 'jitter')+
  scale_color_brewer(type = 'qual', palette = 'Set1',
                     guide = guide_legend(title = 'Alcohol_precent', reverse = T,
                                          override.aes = list(alpha = 1, size = 3)))+
  coord_cartesian(ylim = c(min(pf$fixed.acidity), quantile(pf$fixed.acidity, 0.99)))+
  facet_wrap(~grades, ncol = 4)+
  ggtitle('Grades scatter plot by Density and log(fixed acidity) colored by Alcohol precentage')
```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}
ggplot(pf, aes(citric.acid, pH, color = grades))+
  geom_point(alpha = 0.5, size = 2, position = 'jitter')+
  scale_color_brewer(type = 'div', 
                     guide = guide_legend(title = 'Quality Grades', reverse = T,
                                          override.aes = list(alpha = 1, size = 3)))+
  ggtitle('Relatioship between pH and citric acid coloured by Quality')


```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!